---
title: "stroke"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import Library
```{r}
library(tidyverse)
library(readxl)
library(rms)
```


## Read & Clean Data
```{r cars}
setwd("C:/UBC MBAN Coursework/3 BABS 508 Advanced Predictive Business Analytics/Project")
dat <- read_excel('Stroke.xlsx')
# dat <- read.csv('Stroke.csv')
dim(dat) # Check dimension
dat <- na.omit( dat )
dat <- dat %>% filter(bmi != 'N/A') # Remove rows with 'N/A' bmi value 
dim(dat) # Check dimension after removing missing values
```


## Explore Data
```{r}
d <- describe(dat)
html(d, size = 80, scroll = TRUE)
```


## Factorize/Numerize variables + change reference group for smoke
```{r transform}
dat$gender <- as.factor(dat$gender)
dat$hypertension <- as.factor(dat$hypertension)
dat$heart_disease <- as.factor(dat$heart_disease)
dat$ever_married <- as.factor(dat$ever_married)
dat$work_type <- as.factor(dat$work_type)
dat$Residence_type <- as.factor(dat$Residence_type)
dat$smoking_status <- as.factor(dat$smoking_status)
dat$smoking_status <- relevel(dat$smoking_status, ref = 'never smoked')
dat$bmi <- as.numeric(dat$bmi)
# dat$stroke <-as.factor(dat$stroke)
```



## Two-way (categorical)
```{r seecrosstab}
xtabs(~stroke + gender, data = dat)
xtabs(~stroke + hypertension, data = dat)
xtabs(~stroke + heart_disease, data = dat)
xtabs(~stroke + ever_married, data = dat)
xtabs(~stroke + work_type, data = dat)
xtabs(~stroke + Residence_type, data = dat)
xtabs(~stroke + smoking_status, data = dat)
```


## Prepare the data
```{r selectpred, echo=FALSE}
# List of names of variables to analyze
v <- c('stroke','gender','age','hypertension','heart_disease',
       'ever_married', 'work_type', 'Residence_type', 'avg_glucose_level', 'bmi', 
       'smoking_status')
dat <- dat[, v]
units(dat$age) <- 'years'
```


## Summarize data
```{r summarize, echo=FALSE}
s <- summary(gender + age + hypertension + heart_disease + ever_married + 
             work_type + Residence_type + avg_glucose_level + bmi + smoking_status 
              ~ stroke, data = dat, overall = TRUE)


html(s, caption='Predictors according to stroke(Y/N)',
     exclude1 = TRUE, npct = 'both', digits = 2,
     prmsd = TRUE, brmsd = TRUE, msdsize = mu$smaller2)

```


## Correlation Matrix
```{r seecorrelation}
tmp <- data.matrix(data.frame(unclass(dat))) #label categorical data into numeric
rcorr(tmp)
```



## Visualize the relationship between the continuous variables and the outcome (stroke) to assess linearity

### Histspike bins the continuous x variable into equal-width bins and then computes and plots the frequency counts of Y within each bin. The function then displays the proportions as a vertical histogram with a loess curve fit to the plot.The loess nonparametric smoother is an excellent tool for determining the shape of the relationship between a predictor and the response
```{r seelin, echo=FALSE}
# datadist function computes statistical summaries of predictors to  automate 
# estimation and plotting of effects
dd <- datadist(dat)
options(datadist = "dd")

a <- ggplot(dat, aes(x = age, y = stroke)) +
  histSpikeg(stroke ~ age, lowess = TRUE, data = dat) +
  labs(x = "\nAge", y = "Probability(Stroke)\n")

a


b <- ggplot(dat, aes(x = avg_glucose_level, y = stroke)) +
  histSpikeg(stroke ~ avg_glucose_level, lowess = TRUE, data = dat) +
  labs(x = "\nAvg_glucose_level", y = "Probability(Stroke)\n")

b

c <- ggplot(dat, aes(x = bmi, y = stroke)) +
  histSpikeg(stroke ~ bmi, lowess = TRUE, data = dat) +
  labs(x = "\nbmi", y = "Probability(Stroke)\n")

c
```


## Consider polynomial transformations for possible non-linearity
```{r seepoly1, echo=FALSE}
quad_bmi <- lrm(stroke ~ gender + age + hypertension + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + poly(bmi, 2) + smoking_status , data = dat)
print(quad_bmi)
```

```{r seepoly2, echo=FALSE}
# cube_glucoe <- lrm(stroke ~ gender + age + hypertension + heart_disease + ever_married + work_type + Residence_type + poly(avg_glucose_level,3) + bmi + smoking_status , data = dat)
# print(cube_glucose)
```

## Suspect interactions between the effect of age and other predictor variables...initial examination using Analysis of Variance
```{r aov_age_gender, echo=FALSE}
# Testing interaction between age and gender
m = aov(stroke ~ age * gender, dat)
summary(m)
```

```{r aov_age_ht, echo=FALSE}
# Testing interaction between age and hypertension
m = aov(stroke ~ age * hypertension, dat)
summary(m)
```

```{r aov_age_bmi, echo=FALSE}
# Testing interaction between age and bmi
m = aov(stroke ~ age * bmi, dat)
summary(m)
```

```{r aov_age_ss, echo=FALSE}
# Testing interaction between age and smoking_status
m = aov(stroke ~ age * smoking_status, dat)
summary(m)
```

```{r aov_age_em, echo=FALSE}
# Testing interaction between age and ever_married
m = aov(stroke ~ age * ever_married, dat)
summary(m)
```

```{r aov_age_rt, echo=FALSE}
# Testing interaction between age and Residence_type
m = aov(stroke ~ age * Residence_type, dat)
summary(m)
```

```{r aov_age_hd, echo=FALSE}
# Testing interaction between age and heart_disease
m = aov(stroke ~ age * heart_disease, dat)
summary(m)
```

```{r aov_age_agl, echo=FALSE}
# Testing interaction between age and avg_glucose_level
m = aov(stroke ~ age * avg_glucose_level, dat)
summary(m)
```

```{r aov_age_wt, echo=FALSE}
# Testing interaction between age and work_type
m = aov(stroke ~ age * work_type, dat)
summary(m)
```

```{r multicol, echo=FALSE}
checkint <- lrm(stroke ~ gender + age + hypertension + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + bmi + smoking_status , data = dat, , x=TRUE, y= TRUE)
print(checkint)

# The VIF function in RMS Computes variance inflation factors from the covariance matrix of parameter estimates
# RMS VIF will provide estimates for categorical variables
vif(checkint)
```


##  Suspect interactions between the effect of age and two predictor variables
```{r seeinteraction1, echo=FALSE}
a_and_h <- lrm(stroke ~ gender + age + hypertension + age*hypertension + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + bmi + smoking_status , data = dat)
print(a_and_h)
```

```{r seeinteraction2, echo=FALSE}
a_and_s <- lrm(stroke ~ gender + age + hypertension + age*smoking_status + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + bmi + smoking_status , data = dat)
print(a_and_s)
```


## Stepwise Variable Selection
### perform backward stepwise variable selection - the final model should be adjusted later
```{r bwstep, echo=FALSE}
fastbw(checkint)
```


## Final Model
```{r finalmodel, echo=FALSE}
final_model <- lrm(stroke ~ age + hypertension, data = dat,x = TRUE, y = TRUE)
print(final_model)
```

## Influential observation
```{r finalmodelinfl, echo=FALSE}
u4 <- which.influence (final_model, .4) 
print(u4)
u2 <- which.influence (final_model, .2)  
print(u2)
```

```{r noinfl, echo=FALSE}
x <- c(34, 54, 138, 247, 268, 432, 437)
dat_noinfl <- dat[-x,]
```


```{r withoutinfl, echo=FALSE}
# without influential obs
final_model <- lrm(stroke ~ age + hypertension, data = dat_noinfl,x = TRUE, y = TRUE)
print(final_model)
```


```{r withinfl, echo=FALSE}
# with influential obs
final_model <- lrm(stroke ~ age + hypertension, data = dat,x = TRUE, y = TRUE)
print(final_model)
```


## Model details
### Print probabilities of stroke by age and hypertension, including influential observations
```{r withinfl_prob, echo=FALSE}
# with influential obs
dd <- datadist(dat); options(datadist='dd')
probs <- Predict(final_model, age, hypertension,  fun=plogis)
# probs <- Predict(final_model, age=c(20, 40, 60, 80, 100), hypertension,  fun=plogis)
print(probs)
ggplot(probs) + labs(title="Effects of Age on log-odds of Stroke by Hypertension")
```

### Probability of stroke for team members' age in different hypertesion status
```{r exp_int, echo=FALSE}
# datadist function computes statistical summaries of predictors to  automate estimation and plotting of effects
probs <- Predict(final_model, age=c(22,25,32), hypertension=c(0,1), fun=plogis)
print(probs)

# dlogis gives the density, plogis gives the distribution function, qlogis gives the quantile function, and rlogis generates random deviates.
# The length of the result is determined by n for rlogis, and is the maximum of the lengths of the numerical arguments for the other functions.
```


## Validating the model using the bootstrap to detect whether the model is overfitting the data
```{r val_final, echo=FALSE}
# Update will update and (by default) re-fit a model. 
# x - causes the expanded design matrix (with missings excluded) to be returned under the name x. For print, an object created by lrm.
# y- causes the response variable (with missings excluded) to be returned under the name y.
final_model <- update (final_model, x=TRUE, y=TRUE)  
val <- validate(final_model, B=200, bw=TRUE) 
```

```{r cal_final, echo=FALSE}
cal <- calibrate(final_model, B =200)
plot(cal)
```

```{r val_final_output, echo=FALSE}
# Update will update and (by default) re-fit a model. 
# x - causes the expanded design matrix (with missings excluded) to be returned under the name x. For print, an object created by lrm.
# y- causes the response variable (with missings excluded) to be returned under the name y.
final_model <- update (final_model, x=TRUE, y=TRUE)  
val <- validate(final_model, B=200, bw=TRUE) 
# Printing the variables selected in the first 20  bootstrap resamples shows that the final selected variables are somewhat arbitrary 
print(val, B=50, digits =3) 
```

### You can use the the optimism-corrected slope in the validate output to adjust the original prediction model for optimism
```{r opt_corr, echo=FALSE}
shrinkage.factor <- val["Slope","index.corrected"]
#You multiply the model coefficients (except intercept) by the shrinkage factor (the optimism-corrected slope)
corr_coef <- data.frame(Original = coef(final_model), shrunk.boot = c(coef(final_model)[1], coef(final_model)[-1] * shrinkage.factor))
round(corr_coef, 3)
```



## Alternative Model building
```{r Alternative_model_PCA, echo=FALSE}
# with influential obs

dat.pca <- dat

dat.pca$gender <- as.numeric(dat$gender)
dat.pca$hypertension <- as.numeric(dat$hypertension)
dat.pca$heart_disease <- as.numeric(dat$heart_disease)
dat.pca$ever_married <- as.numeric(dat$ever_married)
dat.pca$work_type <- as.numeric(dat$work_type)
dat.pca$Residence_type <- as.numeric(dat$Residence_type)
dat.pca$smoking_status <- as.numeric(dat$smoking_status)
#dat$smoking_status <- relevel(dat$smoking_status, ref = 'never smoked')
dat.pca$bmi <- as.numeric(dat$bmi)
```

```{r incpca, echo=FALSE}
# Function to compute the first k PCs
ipc <- function(x, k=1, ...)
  princomp(x, ..., cor=TRUE)$scores[,1:k]
# Compute the first 2 PCs
pc2 <- ipc(~ gender + age + hypertension + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + bmi + smoking_status,
           data=dat.pca, k=2)
pc2_mod   <- lrm(stroke ~ pc2, data=dat.pca)

print(pc2_mod)
```

```{r compare, echo=FALSE}
c(final_model=AIC(final_model), pc2_mod=AIC(pc2_mod)) 
```










